---
title: "Final-Project-Data"
format: html
editor: visual
---
# Uploading the data
```{r}
strawberrydata <- read.csv("C:\\Users\\kevin\\OneDrive\\Documentos\\R\\Final Project\\Data\\2024RenovationStrawberryData.csv") 
```
# 
```{r}
head(strawberrydata)
summary(strawberrydata)

```
```{r}
library(dplyr)
```

# Cleaning data for marketable yield to discover each variety have better perfomance for market.
```{r}
data_clean <- strawberrydata %>%
  drop_na(Marketable.Yield..g.)
  
```

# The average yield for each Variety under each Treatment.
```{r}
yield <- data_clean %>%
  group_by(Variety,Treatment) %>%
  summarize(Mean_Yield = mean(Marketable.Yield..g.), .groups = "drop")

kable(yield)
```
```{r}
library("ggplot2")
```


```{r}
ggplot(data = yield, aes(x = Variety, y = Mean_Yield, fill = Treatment)) +
  geom_bar(stat = 'identity', position = "dodge") +
  labs(
    title = "Marketable Crop Yield Comparison by Variety and Mulch type",
    x = "Variety",
    y = "Mean Yield"
  ) +
  scale_fill_manual(values = c("Plastic" = "black","Straw" = "brown"))+
  theme_bw() 
```

# Statistical Model - ANOVA - for Yield
```{r}
anova_model <- aov(Marketable.Yield..g. ~ Variety * Treatment, data = data_clean)

summary(anova_model)

```
# Results: 
### Varieties does not have significant different yields.
### Mulch type affects significatly the yields.
### The mulch type affects equally all strawberry varieties. 

# ANOVA Table:
```{r}
anova_results <- summary(anova_model)

anova_table <- broom::tidy(anova_model) %>%
  select(term, df, statistic, p.value) %>%
  rename(
    Term = term,
    DF = df,
    F_Statistic = statistic,
    P_Value = p.value
  )

print(anova_table)
```

```{r}
install.packages("ggpubr")

```
```{r}
library(ggpubr)
ggplot(anova_model, aes(x = Treatment, y = Marketable.Yield..g., fill = Treatment)) +
  geom_boxplot() +
  stat_summary(fun = "mean", geom = "point", shape = 20, size = 3, color = "black") +
  labs(
    title = "Effect of Mulch Treatment on Yield",
    x = "Treatment",
    y = "Yield"
  ) +
  theme_minimal() +
  scale_fill_manual(values = c("Plastic" = "lightblue", "Straw" = "lightgreen"))
  
```

```{r}
ggplot(anova_model, aes(x = interaction(Variety, Treatment), y = Marketable.Yield..g., fill = Treatment)) +
  geom_boxplot() +
  labs(
    title = "Yield by Variety and Treatment",
    x = "Variety and Treatment",
    y = "Yield"
  ) +
  scale_fill_manual(values = c("Plastic" = "lightblue", "Straw" = "pink")) +
  theme_minimal()

```

```{r}

```


```{r}
ggplot(data = data_clean, aes(x = Variety, y = Marketable.Yield..g., fill = Variety)) +
  geom_boxplot() +
  labs(title = "Distribution of Marketable Yield by Variety",
    x = "Variety",
    y = "Yield"
  ) +
  theme_minimal() 

```

```{r}
ggplot(data = data_clean, aes(x = Treatment, y = Marketable.Yield..g., fill = Treatment)) +
  geom_boxplot() +
  labs(title = "Distribution of Marketable Yield by Treatment",
    x = "Treatment",
    y = "Yield"
  ) +
  scale_fill_manual(values = c("Plastic" = "black","Straw" = "brown"))+
  theme_minimal() 
```


### SPACE


```{r}
vigor <- data_clean %>%
  group_by(Variety,Vigor) %>%
  summarize(Mean_Vigor = mean(Vigor), .groups = "drop")

vigor_nona <- vigor %>%
  drop_na(Vigor)

print(vigor)

```
```{r}
ggplot(data_clean, aes(x = Treatment, y = Vigor, fill = Treatment)) +
  geom_boxplot() +
  labs(
    title = "Comparison of Vigor Scores by Treatment",
    x = "Treatment",
    y = "Vigor Score"
  ) +
  theme_minimal()

```

```{r}
vigor_na <- data_clean %>%
  group_by(Treatment,Vigor)



ggplot(data = vigor_na, aes(x = Treatment, y = Vigor, fill = Treatment)) +
  geom_boxplot() +
  labs(title = "Distribution of Vigor Scores by Variety",
    x = "Variety",
    y = "Vigor Score"
  ) +
  theme_minimal() 

```
```{r}

ggplot(data_clean, aes(x = Vigor, y = Marketable.Yield..g., colour = Treatment)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, aes(group = Treatment), linetype = "dashed") +
  facet_wrap(~ Variety) +  # Facet by Variety
  labs(
    title = "Scatter Plot: Vigor Scores vs. Yield (Faceted by Variety)",
    x = "Vigor Score",
    y = "Yield",
    color = "Mulch Type"
  ) +
  theme_minimal() +
  scale_color_brewer(palette = "Set2")

```


```{r}
ggplot(data = data_clean, aes(x = Variety, y = Marketable.Yield..g., fill = Treatment)) +
  geom_boxplot()+
  labs(
    title = "Marketable Crop Yield Comparison by Variety and Mulch type",
    x = "Variety",
    y = "Yield"
  ) +
  scale_fill_manual(values = c("Plastic" = "black","Straw" = "brown"))+
  theme_bw() 
```




####

# Cleaning data for marketable yield to discover each variety have better perfomance for market.
```{r}
un_data_clean <- data_clean %>%
  drop_na(Unmarketable.Yield..g.)
  
```

# The average yield for each Variety under each Treatment.
```{r}
un_yield <- un_data_clean %>%
  group_by(Variety,Treatment) %>%
  summarize(Mean_un_Yield = mean(Unmarketable.Yield..g.), .groups = "drop")

print(un_yield)
```

```{r}
ggplot(data = un_yield, aes(x = Variety, y = Mean_un_Yield, fill = Treatment)) +
  geom_bar(stat = 'identity', position = "dodge") +
  labs(
    title = "Unmarketable Crop Yield Comparison by Variety and Mulch type",
    x = "Variety",
    y = "Mean Yield"
  ) +
  scale_fill_manual(values = c("Plastic" = "black","Straw" = "brown"))+
  theme_bw() 
```


